A comparison of classification algorithms applied to the Adult Database

### 2009-12-10 report/final.tex

// TODO: Tabela e descrição das diferentes taxas de erro

// TODO: Descrição de alguns padrões interessantes que sobressaiam nos modelos

>>>>

\begin{table}
    \begin{center}
    \begin{tabular}{ | l | l | }
    \hline
    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
    CN2 & 15.20 \\ \hline
    ID3 & 15.68 \\ \hline
    ADTree & 19.80 \\ \hline
    ILP (Aleph) & 20.45 \\ \hline
    Bayesian Networks (Simulated Annealing) & 21.09 \\ \hline
    C4.5 & 21.69 \\ \hline
    Random Forest & 22.71 \\ \hline
    Bayesian Networks (Hill Climbing) & 22.93 \\ \hline
    CART & 25.37 \\ \hline
    Bayesian Networks (Genetic Search) & - \\ \hline
    \end{tabular}
    \caption{Error rates for the different algorithms used in the classification process}
    \label{tbl:overall_results}
    \end{center}
\end{table}

Table \ref{tbl:overall_results} summarizes the best error rates for the 
different algorithms used in the classification process. Each algorithm has 
different parameters to tweak it and the table only shows the best error rates
obtained. CN2 has the best results for this particular dataset, although it
takes longer than the majority of the other algorithms to build a classifier. 
ID3 and C4.5 are usually quick to process and build classifiers within a decent
error rate. ILP and CN2 build more expressive classifiers, capable of a better
interpretation by a human. One special note goes to Inductive Logic Programming,
a technique that allows the use of logic predicates to describe both the 
background knowledge and the produced model. The use of prolog enables the 
output produced by this technique to be better analyzed by a human agent. 
Another important conclusion of the conducted experimental process is that the 
worst error rate of the algorithms used is near 25\%, which denotes the power
of the current classification algorithms, capable of building relevant models in
a short period of time for relatively large datasets.

Analyzing the models produced, one can see that having a high capital gain 
clearly contributes to gaining more than 50K dollars per year. On the other
hand, part time workers usually don't gain more than 50K dollars per year, like
the younger and not married people. Having a higher education clearly 
contributes to gaining more money, as does a stable life (married people
gain more). People who work in private businesses or in the government are also
more suitable to gain above 50K dollars per year. Working more hours per week
obviously contributes to getting more money. Some less obvious conclusions show 
that being a black own child, even though working more than 40 hours per week 
doesn't lead to winning more than 50K dollars per year. In fact, being an own 
child appears on a lot of rules describing people who win less than 50K. ILP
also builds some interesting clauses where the nationality is associated with.
This shows that being a native of some asian countries, such as taiwan, thailand
or cambodia usually lead to gaining more than 50K dollars per year.

Classification is a very active area on the machine learning research field, as
the great amount of software tools and algorithms to tackle it show. It has
arrived to a state where efficient models are built fast and with an expression
suitable for analysis by a human agent. Progresses in relational data mining are
allowing models to infer relations between entities, expanding their range of 
applications by a lot. This work enabled a broad study of different types of 
classification algorithms, analyzed both in terms of efficiency and 
expressiveness of the models produced. Taking that into account, classification
is nowadays a very reliable process for data analysis at large.
